## Q1. Data Types and Variables

```{r q1}
# Import dataset from URL
url <- "https://raw.githubusercontent.com/jbrownlee/Datasets/master/airline-passengers.csv"
airline_data <- read.csv(url)

# Identify data types
str(airline_data)

# Create variables for total rows and columns
total_rows <- nrow(airline_data)
total_cols <- ncol(airline_data)

print(paste("Total Rows:", total_rows))
print(paste("Total Columns:", total_cols))

# Convert numeric column (Passengers) to integer
# Note: It might already be integer, but enforcing it as requested
airline_data$Passengers <- as.integer(airline_data$Passengers)

# Convert character column (Month) to factor
airline_data$Month <- as.factor(airline_data$Month)

# Check types again
class(airline_data$Passengers)
class(airline_data$Month)
```

## Q2. Operators in R

### Part A: Arithmetic Operators - Total Passengers

```{r q2a}
total_passengers <- sum(airline_data$Passengers)
print(paste("Total Passengers:", total_passengers))
```

### Part B: Relational Operators - Months with \> Average Passengers

```{r q2b}
avg_passengers <- mean(airline_data$Passengers)
high_traffic_months <- airline_data[airline_data$Passengers > avg_passengers, ]
head(high_traffic_months)
```

### Part C: Logical Operators - Passengers \> 300 AND \< 500

```{r q2c}
mid_range_traffic <- airline_data[airline_data$Passengers > 300 & airline_data$Passengers < 500, ]
head(mid_range_traffic)
```

### Part D: Assignment Operators

```{r q2d}
# Storing total in a variable (already done in Part A, repeating for clarity)
total_val <- sum(airline_data$Passengers)
print(total_val)
```

## Q3. Control Structures

```{r q3}
# If-else statement to check if average is high or low (arbitrary threshold 250)
if (avg_passengers > 250) {
  print("Average passenger count is High")
} else {
  print("Average passenger count is Low")
}

# For loop to display first 10 values of Passengers
print("First 10 Passenger values:")
for (i in 1:10) {
  print(airline_data$Passengers[i])
}

# While loop to count rows where Passengers > 400
count <- 0
i <- 1
while (i <= nrow(airline_data)) {
  if (airline_data$Passengers[i] > 400) {
    count <- count + 1
  }
  i <- i + 1
}
print(paste("Count of months with > 400 passengers:", count))

# Use ifelse() to classify values
# Categories: Low (< 200), Medium (200-400), High (> 400)
airline_data$Category <- ifelse(airline_data$Passengers < 200, "Low",
                                ifelse(airline_data$Passengers <= 400, "Medium", "High"))
head(airline_data)

# Explanation of if vs ifelse
# 'if' statement is a control structure used for conditional execution of a block of code based on a single condition (scalar).
# 'ifelse()' is a vectorized function that checks a condition for every element of a vector and returns a value depending on whether the element matches the condition.
```

## Q4. Functions in R

```{r q4}
# Function for Mean, Median, and SD
analyze_stats <- function(numeric_col) {
  col_mean <- mean(numeric_col)
  col_median <- median(numeric_col)
  col_sd <- sd(numeric_col)
  
  return(list(Mean = col_mean, Median = col_median, StandardDeviation = col_sd))
}

# Call the function
stats <- analyze_stats(airline_data$Passengers)
print(stats)

# Function to find Max and Min
find_extremes <- function(numeric_col) {
  max_val <- max(numeric_col)
  min_val <- min(numeric_col)
  return(paste("Maximum:", max_val, " | Minimum:", min_val))
}

# Call the function and display output
extremes <- find_extremes(airline_data$Passengers)
print(extremes)
```
