# Part A: Dataset Understanding and Preparation

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
url <- "https://raw.githubusercontent.com/datasets/covid-19/main/data/countries-aggregated.csv"
covid <- read.csv(url)

```

```{r}
head(covid, 10)
tail(covid, 10)

```

```{r}
dim(covid)

```

```{r}
str(covid)

```

```{r}
covid$Date <- as.Date(covid$Date)

```

# Part B: Data Cleaning and Preprocessing

```{r}
colSums(is.na(covid))

```

```{r}
covid_clean <- covid %>%
  drop_na()

```

```{r}
Q1 <- quantile(covid_clean$Confirmed, 0.25)
Q3 <- quantile(covid_clean$Confirmed, 0.75)
IQR_val <- Q3 - Q1

covid_clean <- covid_clean %>%
  filter(Confirmed >= (Q1 - 1.5 * IQR_val),
         Confirmed <= (Q3 + 1.5 * IQR_val))

```

# Part C: Data Manipulation

```{r}
japan_data <- covid_clean %>%
  filter(Country == "Japan")
```

```{r}
japan_2020 <- japan_data %>%
  filter(year(Date) == 2020)

```

```{r}
covid_clean <- covid_clean %>%
  mutate(Active_Cases = Confirmed - Recovered - Deaths)

```

```{r}
country_summary <- covid_clean %>%
  group_by(Country) %>%
  summarise(
    Total_Confirmed = max(Confirmed),
    Total_Deaths = max(Deaths),
    Total_Recovered = max(Recovered)
  )

```

```{r}
top10_confirmed <- country_summary %>%
  arrange(desc(Total_Confirmed)) %>%
  slice(1:10)

```

# Part D: Data Visualization

### Scatter Plot: Confirmed vs Deaths

```{r}
ggplot(covid_clean, aes(x = Confirmed, y = Deaths)) +
  geom_point(alpha = 0.4) +
  labs(title = "Confirmed Cases vs Deaths",
       x = "Confirmed Cases",
       y = "Deaths")

```

# Line Chart: Confirmed Cases Over Time (Japan)

```{r}
ggplot(japan_data, aes(x = Date, y = Confirmed)) +
  geom_line(color = "blue") +
  labs(title = "COVID-19 Confirmed Cases Trend in Japan",
       x = "Date",
       y = "Confirmed Cases")

```

# Comparison of Three Countries

```{r}
compare_countries <- covid_clean %>%
  filter(Country %in% c("Japan", "US", "Brazil"))

ggplot(compare_countries, aes(x = Date, y = Confirmed, color = Country)) +
  geom_line() +
  labs(title = "Confirmed Cases Comparison",
       x = "Date",
       y = "Confirmed Cases")

```

### Bar Chart: Total Deaths (Top 10 Countries)

```{r}
ggplot(top10_confirmed, aes(x = reorder(Country, Total_Deaths), y = Total_Deaths)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Total Deaths by Country (Top 10)",
       x = "Country",
       y = "Deaths")

```

# Horizontal Bar Chart: Recovered Cases

```{r}
ggplot(top10_confirmed, aes(x = Total_Recovered, y = reorder(Country, Total_Recovered))) +
  geom_bar(stat = "identity") +
  labs(title = "Recovered Cases by Country",
       x = "Recovered Cases",
       y = "Country")

```

# Histogram: Confirmed Cases

```{r}
ggplot(covid_clean, aes(x = Confirmed)) +
  geom_histogram(bins = 5, fill = "gray", color = "black") +
  labs(title = "Distribution of Confirmed COVID-19 Cases",
       x = "Confirmed Cases",
       y = "Frequency")

```

**Interpretation:**\

Most countries have relatively low confirmed cases, while a few have very high values, indicating uneven healthcare strain.

# Box Plots

```{r}
ggplot(covid_clean, aes(y = Confirmed)) +
  geom_boxplot() +
  labs(title = "Box Plot of Confirmed Cases")

```

```{r}
ggplot(covid_clean, aes(y = Deaths)) +
  geom_boxplot() +
  labs(title = "Box Plot of Deaths")

```

**Observation:**\
Outliers indicate countries with significantly higher COVID-19 impact compared to others.
